\bf The \texttt{forestecology} R package for modeling interspecies competition between trees

### 2021-02-12 etc/paper/forestecology.tex
Done "compute the growth of trees based on census data" section

\if0\blind
{
  \title{\bf \texttt{forestecology} package for modeling interspecies competition
between trees}

@@

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf \texttt{forestecology} package for modeling interspecies competition
between trees}
  \end{center}
  \medskip
} \fi

@@

\hypertarget{abstract-350-words}{%
\section{Abstract (350 words)}\label{abstract-350-words}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set the context for and purpose of the work: The scientific
  question/problem and the desiderata

  \begin{itemize}
  \tightlist
  \item
    (Eventually) modularly fitting models for interspecific competition
    and assessing them using spatial crossvalidation
  \item
    Leverage ForestGEO protocols providing standardization
  \end{itemize}
\item
  Indicate the approach and the methods: Use tidyverse, simple features,
  and tidymodels packages.

  \begin{itemize}
  \tightlist
  \item
    tidyverse: as stated in tidy tools manifesto: standadized data
    structures, functional programming with pipe, designed for humans
  \item
    sf: tidyverse-friendly package that makes wrangling and visualizing
    spatial data much easier
  \item
    tidymodels: given our spatial-crossvalidation, use tidymodels
    framework is a collection of packages for modeling and machine
    learning using tidyverse principles
  \item
    Dave: hmm, I think here we might need to mention the methods we
    implement. Something like, the package provides functions to
    specific a linear, bayesian neighborhood competition model of
    growth, then fit the model, and compare competing models with
    spatiall cross validation. Or somethign like that.
  \end{itemize}
\item
  Outline the main results: We replicate from scratch the figure in
  PLOSOne paper using Big Woods data, conduct similar analysis for SCBI
  data.

  \begin{itemize}
  \tightlist
  \item
    Code would've been way more complicated in base.
  \item
    We don't have to worry about how the component functions work, only
    that sequence/converyor belt is correct and output is correct.
  \item
    Scientist is abstracted away from ugly programming details.
  \end{itemize}
\item
  Identify the conclusions and wider implications:

  \begin{itemize}
  \tightlist
  \item
    New scientific conclusions from SCBI data
  \item
    Modularly switch out our bayesian lm() functions to anything you
    want.
  \item
    this can serve as blue print for other modeling situations.
  \end{itemize}
\end{enumerate}

@@

@@

We demonstrate the \texttt{forestecology} package's features on two data
sets, both based on inventory censuses of two sites from the Smithsonian
Institution's ForestGEO international network of 72 long‐term forest
dynamics research sites \citet{andersonteixeira_ctfs-forestgeo_2015}.
First, the Michigan Big Woods Forest Dynamics Plot located at the Edwin
S. George Reserve in Pinckney, MI, USA. The 23 ha plot is situated in
mature oak-hickory forest. The canopy is dominated by white oak
(\emph{Quercus alba}), northern red oak (\emph{Q. rubra}), black oak
(\emph{Q. velutina}), shagbark hickory (\emph{Carya ovata}) and pignut
hickory (\emph{C. glabra}) \citet{allen_michigan_2020}. In the example
below, we will preface any data frames from this plot in with
\texttt{bw\_}.

Second, the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot, located at the Smithsonian's National Zoo and
Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640
x 400 m) plot is located at the intersection of three of the major
physiographic provinces of the eastern US: the Blue Ridge, Ridge and
Valley, and Piedmont provinces and is adjacent to the northern end of
Shenandoah National Park. The forest type is typical mature secondary
eastern mixed deciduous forest, with a canopy dominated by tulip poplar
(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
hickories (\emph{Carya} spp.), and an understory composed mainly of
spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}. In the example
below, we will preface any data frames from this plot in with
\texttt{scbi\_}.

The code that generates Figures are included in the supplementary
materials.

We load all the necessary packages.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(forestecology)}
\KeywordTok{library}\NormalTok{(blockCV)}
\end{Highlighting}
\end{Shaded}

\hypertarget{preprocess-census-data}{%
\subsection{Preprocess census data}\label{preprocess-census-data}}

We start by preprocessing the census data for both sites. While
ForestGEO data protocols ensure a high degree of standardization between
site, minor variations still exist
\citet{andersonteixeira_ctfs-forestgeo_2015}. While the Big Woods data
comes pre-loaded in the \texttt{forestecology} package, we load the SCBI
data as they are saved in .csv files in the SCBI-ForestGEO-Data
repository on GitHub
\citet{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. In both
cases, we load the census data as R as ``tibble'' data frames thereby
ensuring a standardized input/output format that can be used across all
\texttt{tidyverse} packages \citet{wickham_welcome_2019}.

Furthermore, we ensure that the different variables have the correct
names, types (\texttt{dbl}, \texttt{data}, \texttt{factor}).

We load census data from 2008 and 2014 saved in the package, then merge
species data (genus, species, linnean classification, family, etc).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw <-}\StringTok{ }\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(species_bw, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-data}{%
\subsubsection{SCBI}\label{scbi-data}}

We load census data from 2008 and 2014 from \texttt{.csv} files saved
from GitHub on November 20, 2020. Furthermore, we perform two additional
pre-processing steps. First, in order to speed up computation for
purposes of this example, we only consider a 9 ha subsection of the 25.6
ha of the SCBI site: \texttt{gx} from 0--300 instead of 0--400 and
\texttt{gy} from 300--600 instead of 0--640. Second, in order to
standardize comparisons between Big Woods and SCBI, we convert the units
of dbh from mm to cm. \footnote{A rule of thumb to ascertain if dbh is
  in mm or cm is to verify if the smallest non-zero and non-missing
  measurement is 1 or 10. If the former, then cm. If the later, then mm.
  This is because ForestGEO protocols state that only trees with dbh
  greater or equal to 1cm should be included in censuses. }

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}

\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{compute-annual-growth}{%
\subsection{Compute annual growth}\label{compute-annual-growth}}

MERGE THIS SUBSECTION with previous?

For each plot we then compute average annual growth between the two
censuses using the \texttt{compute\_growth()} function. This function
takes the two census data frames as well as a character indicating which
variable in both data frames uniquely identifies each stem. This
function returns a single data frame that includes a numerical variable
\texttt{growth} reflecting the average annual dbh growth (in cm) of all
trees alive at both time points. Furthermore, variables that (in theory)
remain unchanged between censuses appear only once, such as location
variables \texttt{gx} and \texttt{gy}; as well as species-related
variables. Variables that should change between censuses are suffixed
with \texttt{1} and \texttt{2} indicating the earlier and later
censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}. Here
the resulting data frames are named with some variation of
\texttt{growth\_df}.

After computing the average annual growth for each tree, we ensure to
convert all variables denote species from type character to factors;
this is to ensure that issues of rare species being accounted for in
both training and test sets in our upcoming cross-validation step (see
Section REF)

In the case of Big Woods data, we first remove all trees that were
re-sprouts in the later (2014) census. Additionally, we have included
two classification of tree species: \texttt{species} and
\texttt{family}. To illustrate model comparison, we test competition
models in which individuals are grouped by species and by family.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_bw <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2014}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{id =} \StringTok{"treeID"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\hypertarget{comparison}{%
\subsubsection{Comparison}\label{comparison}}

DAVE: I THINK THIS FIGURE IS UNNEEDED. COULD BE REMOVED FOR SPACE
REASONS. Figure \ref{fig:growth-histogram} displays histograms comparing
the distribution of average annual growth at both sites. Observe that
average annual growth appears higher at the Big Woods site.

@@

{\centering \includegraphics[width=1\linewidth]{Figures/growth-histogram-1}

@@

\caption{Distribution of average annual growth in DBH for both sites.}\label{fig:growth-histogram}
\end{figure}

@@

\hypertarget{run-spatial-cross-validation}{%
\subsection{Run spatial
cross-validation}\label{run-spatial-cross-validation}}

@@

\hypertarget{big-woods-5}{%
\subsubsection{Big Woods}\label{big-woods-5}}

@@

\hypertarget{big-woods-6}{%
\subsubsection{Big Woods}\label{big-woods-6}}

@@

\hypertarget{scbi-5}{%
\subsubsection{SCBI}\label{scbi-5}}

>>>>

\if0\blind
{
  \title{\bf The \texttt{forestecology} R package for modeling interspecies
competition between trees}

@@

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf The \texttt{forestecology} R package for modeling interspecies
competition between trees}
  \end{center}
  \medskip
} \fi

@@

\hypertarget{abstract-350-words}{%
\section*{Abstract (350 words)}\label{abstract-350-words}}
\addcontentsline{toc}{section}{Abstract (350 words)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  When modeling growth of trees forest ecologists often incorporate the
  effect of interspecies competition. Many such models are based on a
  neighborhood effect assumption whereby all trees within a fixed
  distance of all focal trees are considered competitors. Methods are
  needed to evaluate the effect of interspecies competition and to
  assess their quality.
\item
  We present the \texttt{forestecology} package providing methods for
  both 1) evaluating the out-of-sample performance of our model using
  spatial-crossvaliation and 2) testing a null hypothesis of of no
  impact of competitor species' identity on the growth of trees using a
  permutation test. We implement a class and methods using R's S3
  object-oriented system, for a specific linear, Bayesian neighborhood
  competition model of tree growth.
\item
  We demonstrate the package's functions using data from the Smithsonian
  Conservation Biology Institute's large forest dynamics plot, part of
  the ForestGEO network of reseach sites. Given ForestGEO's data
  collection protocols and data formatting standards, the package
  cross-compatibility of code. We show both that 1) competitor species
  identity matters and 2) that not spatially cross-validating leads to
  error estimates that are overly optimistic.
\item
  The package follows \texttt{tidyverse}-like structure whereby
  verb-named functions can be modularly ``piped'' in sequence to
  intuitively display the sequence of steps of analysis from start to
  finish. Additionally, most inputs/outputs of functions assume an are
  of \texttt{sf} class from the simple features package, thereby
  facilitating all wrangling and visualization of geospatial data.
  Lastly, even though our package is currently limited to one specific
  model, the package is setup such that it can be easily extended to
  other models.
\end{enumerate}

@@

The \texttt{forestecology} is designed with ``tidy'' data principles in
mind as \citet{wickham_welcome_2019}.

Given that our data is of geo-spatial nature, we represent our data
using the ``simple features'' \texttt{sf} package class of objects
\citet{pebesma_simple_2018} whereby. While previously the \texttt{sp}
package serves such purposes \citet{pebesma_sp_2005}, the \texttt{sf}
package is designed to interface with the \texttt{tidyverse} suite of
packages.

@@

We demonstrate the \texttt{forestecology} package's features on the
Smithsonian Conservation Biology Institute (SCBI) large forest dynamics
plot, located at the Smithsonian's National Zoo and Conservation Biology
Institute in Front Royal, VA, USA. The 25.6 ha (640 x 400 m) plot is
located at the intersection of three of the major physiographic
provinces of the eastern US: the Blue Ridge, Ridge and Valley, and
Piedmont provinces and is adjacent to the northern end of Shenandoah
National Park. The forest type is typical mature secondary eastern mixed
deciduous forest, with a canopy dominated by tulip poplar
(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
hickories (\emph{Carya} spp.), and an understory composed mainly of
spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}.

A high-level overview of the steps of our analysis pipeline is as
follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compute the growth of trees based on census data
\item
  Add spatial information:

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    Define buffer region trees.
  \item
    Add spatial cross-validation block information.
  \end{enumerate}
\item
  Compute focal versus competitor tree information.
\item
  Fit model and make predictions.
\item
  Additionally: Evaluate model performance using spatial
  cross-validation.
\item
  Additionally: Evaluate the effect of competitor species identity using
  permutation tests.
\end{enumerate}

We load all necessary packages.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(forestecology)}
\end{Highlighting}
\end{Shaded}

\hypertarget{compute-the-growth-of-trees-based-on-census-data}{%
\subsection{Compute the growth of trees based on census
data}\label{compute-the-growth-of-trees-based-on-census-data}}

The first step in the our analysis sequence is to compute the growth of
trees using data from two censuses. The \texttt{compute\_growth()}
function computes growth assuming census data that follows ForestGEO
standards. Despite such standards, minor variations will still exist
between sites thereby necessitating some data wrangling and checking.
For example, the SCBI site records all dbh's in millimeters, whereas the
Michigan Big Woods site records them in centimeters
\citet{andersonteixeira_ctfs-forestgeo_2015}
\citet{allen_michigan_2020}. The data format of other sites may be such
that our \texttt{compute\_growth()} function doesn't work at all.
However, in the end all that matters is that the growth of all trees is
saved in a data frame of class \texttt{sf} whereby the geolocation of
each tree is presented in a \texttt{geometry} variable of type
\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
contains the following variables: a variable uniquely identifying each
tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the dbh
at earlier and later census, and \texttt{growth} of type \texttt{dbl}
double quantifying the average annual growth in centimeters.

We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
existed on GitHub on November 20, 2020. After selecting only relevant
variables, we perform a few additional data wrangling steps: convert the
variable with the date of measurement to be of type \texttt{date},
convert dbh to be in centimeters\footnote{A rule of thumb to determine
  the units of dbh is check if the smallest non-zero and non-missing
  measurement is 1 or 10. If the former, then centimeters. If the later,
  then millimeters. This is because ForestGEO protocols state that only
  trees with dbh greater or equal to 1cm should be included in censuses.},
convert the \texttt{sp} variable containing species information from
type \texttt{chr} character to \texttt{fct} factor (we will discuss the
need for this in Section \ref{spatial-cross-validation}). Furthermore,
in order to speed up computation for purposes of this examplt, we only
consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
from 0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
0--640.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}

\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

These two data frames are then used as the two primary arguments to the
\texttt{compute\_growth()} function, along with the \texttt{id} argument
whereby the user specifies the name of the variable that uniquely
identifies each tree-stem under consideration (note this does not
include resprouts in the later census):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 299.9 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##    stemID sp     dbh1 codes1 status  dbh2 codes2  growth     geometry}
\CommentTok{##     <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>    <dbl>      <POINT>}
\CommentTok{##  1      4 nysy  13.6  M      A       14.2 M       0.103  (14.2 428.5)}
\CommentTok{##  2      5 havi   8.8  M      A        9.6 M;P     0.150   (9.4 436.4)}
\CommentTok{##  3      6 havi   3.25 NULL   A        4   M       0.140     (1.3 434)}
\CommentTok{##  4     77 qual  65.2  M      A       66   M       0.141  (34.7 307.2)}
\CommentTok{##  5     79 tiam  47.7  M      A       46.8 M      -0.161    (40 381.1)}
\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
\CommentTok{## # ... with 7,944 more rows}
\end{Highlighting}
\end{Shaded}

The output \texttt{growth\_scbi} is a single data frame of class
\texttt{sf} that includes a numerical \texttt{growth} reflecting the
average annual growth in dbh (in cm) for all trees that were alive at
both time points as well a \texttt{geometry} variable encoding each
tree's geolocation. Furthermore, variables that (in theory) remain
unchanged between censuses appear only once, such as location variables
\texttt{gx} and \texttt{gy}; as well as species-related variables.
Variables that should change between censuses are suffixed with
\texttt{1} and \texttt{2} indicating the earlier and later censuses,
such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.

Given that \texttt{growth\_scbi} is of class \texttt{sf}, it can be
easily plotted in \texttt{ggplot2} using the \texttt{geom\_sf()}
geometry as seen in Figure \ref{fig:scbi-trees}.

TODO: Rescale points in this plot:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{scale_size}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{range =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-trees-1}

@@

\caption{Growth of trees at SCBI.}\label{fig:scbi-trees}
\end{figure}

@@

\hypertarget{spatial-cross-validation}{%
\subsection{Run spatial
cross-validation}\label{spatial-cross-validation}}

@@

@@

@@
